---
title: "R Notebook"
author: "Vivek Kumar"
date: "null"
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

# Sensitization data analysis from JCMS for Kumar Lab

Vivek Kumar

January 2023
 
# Downloading data from JCMS instructions go here.
@tom and @marina

Add info about directory structure @vivek



```{r}
#library(readr)
library(stringr)
library(dplyr)
#library(reshape2)
library(ggplot2)
library(readxl)
library(lubridate)
library(data.table)
library(Rmisc)
library(tidyfast)
library(Hmisc)

################# SET WORKING DIRECTORY
wd <- "/Users/vkumar/Box Sync/LAB/PSY todo/2024-01-31 JCMS Sens Analysis/Data/"
setwd(wd)

################# READ INPUT FILE 
filename <- "/Users/vkumar/Box Sync/LAB/PSY todo/2024-01-31 JCMS Sens Analysis/Data/TOM_PSYQuery2 (3).xlsx"

# Marina wanted the MouseID to maintain 0 in the front --> upload 012345 which is converted to 12345 with the statement below. It convets MouseID to int
# d1 <- fread(filename, sep = '\t')

# read and maintian the MouseID as character
# d1 <- fread(filename, sep = '\t', colClasses = c("MouseID" = "character"))

d1 <- as.data.table(read_xlsx(filename, sheet = "TOM_PSYQuery2" )) #made this correction to convert to data.table.

################# CREATE A FOLDER FOR OUTPUT FILES 
#  Create an empty folder in directory that is one level higher than the input file

dir.create(file.path("../PSY_Processed"), showWarnings = FALSE)

```
## Add test numbers

```{r}
# Making the date correct, if the date is not formatted correctly, adjust this I changed "%m/%d/%Y"
# d2$TestDate <- mdy_hms(d2$TestDate)  

################# SORT BY MouseID AND TestDate
d2 <- d1[order(MouseID, TestDate)]

################# ADD TEST NUMBER AS A NEW COLUMN
d2 <- d2[ , TestNumber := order(TestDate), by = MouseID]



```


# Filtering data to find the right animals.
Filtering done by 
1. Dose 
2. Tester
3. Strain
4. Specific

Problem animal
AnimalName: 021205_B6NJ−Macrod2_Cyfip2_S968F_F_21894
MouseID: 021205

```{r}
################# FILTER DATA AND PROCESS DATA
################# CHANGE THESE AS NEEDED 

################# FILTER BY DOSE ###################################
Dose <- 10 # this is needed becuase I use Dose as a varialbel. Even if you don't use dose, keep this in place.
d2 <- d2[DrugDose == Dose]

# Dose <- c(7.5, 10, 15)
# d2 <- d1[d1$DrugDose %in% Dose, ]


#################################################################### 


################# FILTER BY TESTER #################################
# d2 <- d2[ExptDoneBy == 'Marina']

# d2 <- d2[ExptDoneBy %in% .('Marina', 'Tom', 'Tom_Marina')]


# this can be a list -- example
# d2 <- d1[ExptDoneBy %in% .('Sandeep','Marina')]

################# FILTER BY STRAIN #################################
# d2 <- d2[strainName %in% .('C57BL/6NJ','C57BL/6J')]

# CONTROLS
# d2 <- d2[strainName %in% .("C57BL/6NJ",
#                           "C57BL/6J",
#                           "B6J-Cyfip2 S968F<N> (GET202)",
#                           "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Bcl11b
# d2 <- d2[strainName %in% .("B6NJ-Bcl11b+/- Cyfip2-S968F<J> H Lethal",
#                           "B6NJ-Bcl11b+/+ Cyfip2-S968F<J> ctrl",
#                           "B6NJ-Cyfip2 S968F<J> (GET204)" )]
# 
# Cxxc4
# d2 <- d2[strainName %in% .("B6NJ-Cxxc4-/- Cyfip2-S968F<J> Hom Poor Breeders",
#                           "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Fndc4 - not working Fndc4 not filtered
# d2 <- d2[strainName %in% .("B6NJ-Cyfip2 S968F<J> (GET204)",
#                            "B6NJ-Fndc4-/- Cyfip2-S968F<J> Hom Breed Well")]

# # Isca1tm1a
# d2 <- d2[strainName %in% .("B6NJ-Isca1tm1a+/- Cyfip2-S968F<J> Hlethal",
#                            "B6NJ-Isca1tm1b+/- Cyfip2-S968F<J> H Lethal",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Kcnd3
# d2 <- d2[strainName %in% .("B6NJ-Kcnd3-/- Cyfip2-S968F<J> Hom Breed Well",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# KCNH3
# d2 <- d2[strainName %in% .("B6NJ-KCNH3-/- Cyfip2-S968F<J> Hom Breed Well",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Macrod2
d2 <- d2[strainName %in% .("B6NJ-Macrod2-/- Cyfip2-S968F<J> Hom Breed Well",
                         "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Pcyox1l
# d2 <- d2[strainName %in% .("B6NJ-Pcyox1l-/- Cyfip2-S968F<J> HOK",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Ppfia3
# d2 <- d2[strainName %in% .("B6NJ-Ppfia3-/- Cyfip2-S968F<J> Hom Poor Breeders",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Ptpru
# d2 <- d2[strainName %in% .("B6NJ-Ptpru-/- Cyfip2-S968F<J> Hom Breed Well",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Rimklb
# d2 <- d2[strainName %in% .("B6NJ-Rimklb-/- Cyfip2-S968F<J> HOK",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Scamp5
# d2 <- d2[strainName %in% .("B6NJ-Scamp5+/- Cyfip2-S968F<J> HLethal",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Slc26a10
# d2 <- d2[strainName %in% .("B6NJ-Slc26a10-/- Cyfip2-S968F<J> Hom Breed Well",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Strn3
# d2 <- d2[strainName %in% .("B6NJ-Strn3+/- Cyfip2-S968F<J> H Lethal",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Stx16
# d2 <- d2[strainName %in% .("B6NJ-Stx16-/- Cyfip2-S968F<J> Hom Poor Breeders",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Tars3
# d2 <- d2[strainName %in% .("B6NJ-Tars3-/- Cyfip2-S968F<J> HOK",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Tmem151b
# d2 <- d2[strainName %in% .("B6NJ-Tmem151b-/- Cyfip2-S968F<J> Hom Breed Well",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]

# Tppp
# d2 <- d2[strainName %in% .("B6NJ-Tppp-/- Cyfip2-S968F<J> HOK",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]


# Zfp536
# d2 <- d2[strainName %in% .("B6NJ-Zfp536+/- Cyfip2-S968F<J> H Lethal",
#                          "B6NJ-Cyfip2 S968F<J> (GET204)" )]


# d2 <- d2[strainName %in% .(
# "(B6J x B6NJ)F1-Cyfip2 S968F<J>", "B6J-Cyfip2 S968F<N> (GET202)", 
# "B6J-Srgap3<S642P> (WT357 GET2924 KI CRISPR)", "B6J-Srgap3<S642P> Cyfip2 S968F<N>  (WT372)", 
# "B6NJ-Bcl11b+/- Cyfip2-S968F<J> HLethal", "B6NJ-Bcl11b+/+ Cyfip2-S968F<J> ctrl", 
# "B6NJ-Cyfip2 S968F<J> (GET204)", "B6NJ-Fndc4-/- Cyfip2-S968F<J> HBreed", 
# "B6NJ-Macrod2-/- Cyfip2-S968F<J> HBreed", "B6NJ-Tmem151b-/- Cyfip2-S968F<J> HBreed", 
# "C57BL/6J", "C57BL/6NJ")
# ]

################# FILTER BY TEST DATE #################################

# d1$TestDate <- mdy_hms(d1$TestDate) ## NO
# d1$TestDate <- ymd(d1$TestDate) ## NO

# Analyze data from 2023

# d2$TestDate <- ymd(d2$TestDate)
# d2 <- d2[year(TestDate) == 2023]
# d2 <- d2[year(TestDate) %between% c(2023, 2024)]

# to filter by month and year
# d2 <- d1[month(d1$TestDate) %in% 4:6 & year(d1$TestDate) %in% 2023]


################# FILTER BY SPECIFIC MICE #################################
# 2023 tests
# for B6J and B6NJ only

# # B6J, B6NJ, GET202, SRGAP3
# MouseID_to_analyze <- c(
# "007013",
# "007014",
# "007015",
# "007016",
# "007017",
# "007018",
# "007071",
# "019656"
# )
# 




```




## Use this code to determine which animals have not been tested 13 times, or have been tested more than 13 times. 

Outputs - 
1. "Filtered_PSY_Data_d2file.tsv" - d2 file. this is filtered file from above. Its been filtered for dose, tester, strain. It has NOT been filtered for number of tests or for number of tests. 

*This part is important.* There are cases of mice with duplicated data.
Use the following two table that are produced to clean data.

2. "StrainFrequencyTable.csv" - this will have a frequency table for each
strain

3. "Number_tests.csv" - this table has one row for each mouse the number of
times its been tested.

**IMPORTANT TO LOOK AT THESE TABLES AND DETERMINE IF ALL DATA IS ACCURATE** 


```{r, warning=FALSE}

################# CREATE OUTPUTS FOR SANITY CHECKS
# DATA WILL BE IN THE PSY_Processed DIRECTORY
setwd(wd)
setwd("../PSY_Processed")

################# WRITE d2
fwrite(d2, file = "Filtered_PSY_Data_d2file.tsv", sep = "\t")

################# WRITE FREQUENCY OF TESTS 
# This can be used to find animals that have not not been tested 13 times for instance.

# Frewuency table that counts will list MouseID and frequency

freqTable <- d2[, .N ,by = .(MouseID, strainName)] 
freqTable <- freqTable[order(strainName, N)] #orders the table above

# condenced table by strain, also renames N to numberTests

StrainFreqTable <- freqTable[, .N, by = .(strainName, numberTests = N)]  
fwrite(StrainFreqTable, "StrainFrequencyTable.csv")

# create and write a table by MouseID to determine which animals are missing data 

fwrite(d2[, .N ,by = .(MouseID, strainName)], 
       file = "Number_tests.csv", 
       sep = ",")


```
## This chunk will filter rows that need to be removed.
This can be edited to include other columns and texts.
@marina has placed "Do Not Use" in column d13. I will use this to filter the data.
operate on the d3 object. d3 is used in the next chunk also. 

Output "AnimalsBadTestsRemoved.csv" table will have animals after the "Do Not Use" is removed.

```{r}


TextToRemove <- c("Do Not Use")

d2 <- d2[!grepl(TextToRemove, d13)]

# write a table of animals after removing bad tests
setwd(wd)
setwd("../PSY_Processed")
fwrite(d2, "AnimalsAfterBadTestsRemoved.csv", sep = ",")


```
## Filter data to find the set of animals to analyze AND Sanity Check Calculations and Plots

This part will only operate on mice that have been tested a certain
number of times.

A completed Sens series should have 13 total tests.

need to remove all empty rows still

FileOutputs (usually csv files)

"d3" is a filter of "d2" that has 13 tests. 

1. "AnimalsWith13Tests.csv" is an output of d3

2. "LongFormatOf_d3.csv" - is the long format of "d3"

3. "ByAnimalCompiled_PSYdata*.pdf" - Minute by minute data for each animal. All test dates are plotted on one page. I print the comment for each group here. 

4. "ByAnimalCompiled_PSYdata_B_*.pdf" - Minute by minute data for each animal and each date. One trace per page. So a large file. I also print the comments for each animal here. 

Comments are scattered in 4 columns. I print them all.
toPlot3$IndividualComments
toPlot3$d14
toPlot3$d13
toPlot1$GroupComments


```{r}

################# PLOT ALL TRACES FOR EACH ANIMAL

################# Only work with animals that have been tested a certain number of times

numberTestsNeeded <- 10 # required number of tests

MiceToKeep <- d2[, .N, by = MouseID][, MouseID[N >= numberTestsNeeded]]

# filter the mice with at 10 tests
d3 <- d2[MouseID %in% MiceToKeep]


# rearrange columns 
### I am rearranging the columns. move all minute by minute data to the end
d3 <- d3 %>% relocate(num_range("",0:120), .after = TestNumber)
d3 <- d3 %>% relocate(num_range("-",60:1), .after = TestNumber)

################# WRITE FREQUENCY OF TESTS 
setwd(wd)
setwd("../PSY_Processed")

# This can be used to find animals that have not not been tested 13 times for instance.

# Frequency table that counts will list MouseID and frequency

freqTable_afterFilter <- d3[, .N ,by = .(MouseID, strainName)] 
freqTable_afterFilter <- freqTable_afterFilter[order(strainName, N)] #orders the table above

# condenced table by strain, also renames N to numberTests

StrainFreqTable_afterFilter <- freqTable_afterFilter[, .N, by = .(strainName, numberTests = N)]  
fwrite(StrainFreqTable_afterFilter, "StrainFrequencyTable_afterFilter.csv")

# create and write a table by MouseID to determine which animals are missing data 

fwrite(d3[, .N ,by = .(MouseID, strainName)], 
       file = "Number_tests_afterFilter.csv", 
       sep = ",")

################# ################# 

# write a table of animals with 10 tests
fwrite(d3, "AnimalsWith10Tests.csv", sep = ",")

# make a long table for plotting
# IF THE NUMBER OF COLUMNS CHANGE, THIS MELT CODE NEEDS TO BE ALTERED. LOOKIN THE "AnimalsWith13Tests.csv" TO DETERMINE HOW MANY COLUMNS ARE BEFORE THE -60 COLUMN.
file4 <- melt(d3, id = 1:70)
fwrite(file4, "LongFormatOf_d3.csv", sep = ",")

# Making the date correct, if the date is not formatted correctly, adjust this I changed "%m/%d/%Y"
# file4$TestDate <- mdy_hms(file4$TestDate)  

# as.integer causes some weird oscillation in data. as.numeric is the way to go.  
file4$variable <- as.numeric(as.character(file4$variable))

##########################################################
### Plot By AnimalID - one plot for each animal
##########################################################

toPlotAnimalNames <- levels(as.factor(file4$AnimalName))

outFileNamePDF <- paste("../PSY_Processed/ByAnimalCompiled_PSYdata",format(Sys.time(),"%Y-%m-%d(%H_%M%p)"),".pdf", sep = "") ## .. means one directory level up

pdf(outFileNamePDF, 8,4)

for (toPlot in toPlotAnimalNames) {
# extract the animal to plot
  toPlot1 <- file4[file4$AnimalName == toPlot,]

# plot the animal
  p1 <- ggplot(data = toPlot1, aes(x = variable, y = value, group = TestDate, color = as.factor(TestDate))) +
    geom_line() + 
    annotate("text", x = 0, y  = 90, label= toPlot1$GroupComments, color = "red") +
    coord_cartesian(ylim = c(1,100)) +
    geom_vline(xintercept = 0, color = 'blue', linewidth = 0.2) +
    ggtitle(toPlot) 

  print(p1)
}

dev.off()

##########################################################
### Plot By AnimalID - one plot for each test
##########################################################

outFileNamePDF <- paste("../PSY_Processed/ByAnimalCompiled_PSYdata_B_",format(Sys.time(),"%Y-%m-%d(%H_%M%p)"),".pdf", sep = "") ## .. means one directory level up

toPlotAnimalNames <- levels(as.factor(file4$AnimalName))

pdf(outFileNamePDF, 4,2)

for (toPlot in toPlotAnimalNames) {
# extract the animal to plot
  toPlot1 <- file4[file4$AnimalName == toPlot,]

  testDates <- levels(as.factor(toPlot1$TestDate))

for (toPlot2 in testDates) {
    toPlot3 <- toPlot1[as.factor(toPlot1$TestDate) == toPlot2,]

# plot the animal
  p1 <- ggplot(data = toPlot3, aes (x = variable, y = value)) +
    geom_line() + annotate("text", x = 0, y  = 90, label= toPlot3$IndividualComments, color = "red", size = 3) +
    geom_line() + annotate("text", x = 0, y  = 80, label= toPlot3$d14, color = "red", size = 3) +
    geom_line() + annotate("text", x = 0, y  = 70, label= toPlot3$d13, color = "red", size = 3) +
    coord_cartesian(ylim = c(1,100)) +
    geom_vline(xintercept = 0, color = 'blue', linewidth = .2) +
    ggtitle(paste(toPlot, toPlot2))

  print(p1)
  }
  }

dev.off()

```




## Calculate minute by minute average of activity for each Strain

This section will produce the following summary files. They can be used in Excel for further analysis. 

CSV files

1.    "Min_SummaryPlots*.csv" - this has strain, test number, minute by minute avg 

2.    SEM_Strain_TestNo*.csv - this has strain, test number, minute by minute SEM data

Plots

"Min_SummaryPlots\*.pdf" - Has 2 plots.

1. minute by minute for each day of sens for each strain 

2. minute by minute for each animal faceted by strain.

## SOMETHING WEIRD - CHECK THAT MANY OF THE 10MG ON DAY 11 (SALINE) HAVE VERY GOOD RESPONSE. THESE ARE PROBABLY AN ERROR
```{r}
setwd(wd)

##################### EXPORT SUMMARY TABLES

# these are column names, from minutes -60 to 120 will be used in calculations. 
# This can be changed as needed for other columns. I 'm using a trick to get a sequence of numbers.
# We could easily use c("column1", "Column 10")

cols <- as.character(seq(from = -60, to = 120, by = 1))

# MEAN
d5 <- d3[,lapply(.SD, mean, na.rm=T), by = .(strainName, DrugDose, TestNumber),
.SDcols = cols] []

# SEM
d6 <- d3[,lapply(.SD, function(x) sd(x, na.rm=T)/sqrt(length(na.omit(x)))), by = .(strainName, DrugDose, TestNumber),
.SDcols = cols] []

# this code pastes the dose in the file name. 
# CAREFUL - if DOSE is a list, it will produce an error.
# if there is an error here, check what Dose is. 
fwrite(d5,paste("../PSY_Processed/Mean_Strain_TestNo_",Dose, format(Sys.time(),"mg_%Y-%m-%d(%H_%M%p)"),".csv", sep = ""))
fwrite(d6,paste("../PSY_Processed/SEM_Strain_TestNo_",Dose, format(Sys.time(),"mg_%Y-%m-%d(%H_%M%p)"),".csv", sep = ""))

 
##################### PLOTS minute by minute by strain
p2 <- ggplot(data = file4, mapping = aes(x = variable, y = value, 
                                         fill = as.factor(strainName),
                                         color = as.factor(strainName))) + 
  theme_bw() +
  stat_summary(fun.data = mean_se, geom = "ribbon", 
               alpha = 0.3) +
  stat_summary(fun = mean, geom="line", color = "grey50", linewidth = 0.2) +
  facet_grid(TestNumber ~ DrugDose) +
  xlim(-60, 60) +
  geom_vline(xintercept = 0, color = 'red', linewidth = .2) + 
  theme(legend.position = "top") + theme(legend.title=element_blank()) + # removes legend title
  labs(x = "Time (Min)", y = "Distance (cm)",   title = (paste("Sens 7.5, 10, 15 mg/kg"))) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Test Number (1-4 Saline | 5-10 Cocaine | 11 Saline | 12, 13 Cocaine)", breaks = NULL, labels = NULL)) +
  scale_x_continuous(sec.axis = sec_axis(~ ., name = "Drug Doses (mg/kg)", labels = NULL, breaks = NULL))

p2

# PLOT minute by minute by each animals
p3 <- ggplot(data = file4, mapping = aes(x = variable, y = value, color = as.factor(strainName), fill = as.factor(strainName), group=AnimalName)) + 
  theme_bw() +
  geom_line(linewidth = 0.2) +
  facet_grid(TestNumber ~ strainName + DrugDose) +
  xlim(-60, 60) + 
  geom_vline(xintercept = 0, color = 'red', linewidth = 0.2) +
  ggtitle(paste("Sens 10mg/kg - Per Animal")) +
  theme(legend.position = "top") + theme(legend.title=element_blank()) + # removes legend title
  labs(x = "Time (Min)", y = "Distance (cm)") +
  scale_y_continuous(limits = c(0,30), sec.axis = sec_axis(~ . , name = "Test Number (1-4 Saline | 5-10 Cocaine | 11 Saline | 12, 13 Cocaine)", breaks = NULL, labels = NULL)) 
# using secondary axis option from ggplot2 "https://stackoverflow.com/questions/11353287/how-do-you-add-a-general-label-to-facets-in-ggplot2"

p3

outFileNamePDF <- paste("../PSY_Processed/Min_SummaryPlots",format(Sys.time(),"%Y-%m-%d(%H_%M%p)"),".pdf", sep = "") 
## .. means one directory level up
pdf(outFileNamePDF, 4, 10)

print(p2)
print(p3)

dev.off()

```

## Plot the Sens Data.

I'm recycling code from previous Sens Analysis. Using File d3 from above
chunk Plots will appear in "Sensitization Analysis\*.pdf" file

```{r}
setwd(wd)

######################################################################
# START PDF STATEMENT
# 
#######################################################################
outFileNamePDF <- paste("../PSY_Processed/Sensitization Analysis-",format(Sys.time(),"%Y-%m-%d(%H_%M%p)"),".pdf", sep = "") 
## .. means one directory level up
pdf(outFileNamePDF, 10,8)

######################################################################
# sum5minavg 
# 
#######################################################################
# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum5minavg, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  theme_bw() +
  theme(legend.position = "top", 
        legend.title = element_blank(),
        plot.margin = margin(t = 2,  # Top margin
                             r = 2,  # Right margin
                             b = 2,  # Bottom margin
                             l = 2,  # Left margin
                             unit = "cm")) +
  facet_grid(.~DrugDose)
   

print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum5minavg, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(.~DrugDose)

print(sp1)



# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum5minavg, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = sum5minavg, colour = as.factor(AnimalName)), data=d3, shape = 21,size = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = sum5minavg, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = sum5minavg, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = sum5minavg, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)

######################################################################
# sum60min 
# 
#######################################################################

# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum60min, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")+
  facet_grid(.~DrugDose)

print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum60min, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(.~DrugDose)

print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum60min, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom")+ theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = sum60min, colour = as.factor(AnimalName)), data=d3, shape = 21,linewidth = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = sum60min, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = sum60min, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = sum60min, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)


######################################################################
# NetResponse 
# 
#######################################################################

# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = NetResponse, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")+
  facet_grid(.~DrugDose)
print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = NetResponse, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")+
  facet_grid(.~DrugDose)
print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = NetResponse, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = NetResponse, colour = as.factor(AnimalName)), data=d3, shape = 21,size = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = NetResponse, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = NetResponse, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = NetResponse, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)

######################################################################
# AvgResp30min 
# 
#######################################################################

# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")+
  facet_grid(.~DrugDose)
print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")+
  facet_grid(.~DrugDose)
print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = as.factor(AnimalName)), data=d3, shape = 21,size = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = AvgResp30min, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = AvgResp30min, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)

######################################################################
# BaselineAvg 
# 
#######################################################################

# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")+
  facet_grid(.~DrugDose)
print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")+
  facet_grid(.~DrugDose)
print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = as.factor(AnimalName)), data=d3, shape = 21,size = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = BaselineAvg, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = BaselineAvg, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = DrugDose ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)

######################################################################
# END PDF STATEMENT
# 
dev.off()
#######################################################################

```

# Data QC - determine why things may be behaving weirdly.

```{r}

# PLOT FACET BY STRAIN, POINT FOR EACH TEST Color by test date
sp1 <- ggplot(aes(x = as.factor(strainName),y = sum5minavg, colour = TestDate,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  geom_line(linewidth = 1.0, stat = "summary") +
  facet_wrap(facets = DrugDose ~ TestNumber) 
print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST Color by test date
sp1 <- ggplot(aes(x = as.factor(TestDate),y = sum5minavg, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  geom_line(linewidth = 1.0, stat = "summary") +
  facet_wrap(facets = DrugDose~ TestNumber) 
print(sp1)



```
## plot sensitization for each animal 

```{r}
##########################################################
### Plot SENS By AnimalID - one plot for each Animal
##########################################################
setwd(wd)


outFileNamePDF <- paste("../PSY_Processed/SENS_by_Animal",format(Sys.time(),"%Y-%m-%d(%H_%M%p)"),".pdf", sep = "") ## .. means one directory level up
# toPlotAnimalNames <- levels(as.factor(d3$AnimalName))
toPlotAnimalNames <- levels(as.factor(d3$MouseID))


pdf(outFileNamePDF, 6,3)

for (toPlot in toPlotAnimalNames) {
  # extract the animal to plot
  toPlot1 <- d3[MouseID == toPlot, ]
  # testDates <- levels(as.factor(toPlot1$TestNumber))

  # select and melt the data columns
  toPlot2 <- melt(toPlot1, id.vars = c("TestNumber"),
                measure.vars = c("sum5minavg",	"sum60min",	"BaselineAvg", "AvgResp30min",	"NetResponse"))

# plot the animal
  p1 <- ggplot() +
  theme_bw() + theme(legend.title=element_blank()) + 
  ggtitle(label = paste("MouseID: ", toPlot), subtitle = paste0("AnimalName: ",levels(as.factor(toPlot1$AnimalName)), "\n", "StrainName: ", levels(as.factor(toPlot1$strainName)))) + 
  geom_point(aes(x = TestNumber,y = value, color = variable), data=toPlot2,size = 3) + 
  geom_line(aes(x = TestNumber,y = value, group = variable, color = variable), data=toPlot2, linewidth = 1.0) +
  scale_y_continuous(limits = c(0,300)) +
  scale_x_discrete(name ="Dose (mg)", limits=c("1","2","3", "4", "5","6", "7","8","9","10", "11", "12", "13")) +
  geom_vline(xintercept=11, color = "blue")

  print(p1)

  }

dev.off()

```


data.table notes <https://rdatatable.gitlab.io/data.table/index.html>

good tutorial
<https://raw.githack.com/uo-ec510-2020-spring/lectures/master/05-datatable/05-datatable.html#1>
