---
title: "R Notebook"
author: "Vivek Kumar"
date: "null"
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

# Sensitization data analysis from JCMS for Kumar Lab

Vivek Kumar

January 2023
 
# Downloading data from JCMS instructions go here.
@tom and @marina

Add info about directory structure @vivek



```{r}
#library(readr)
library(stringr)
library(dplyr)
#library(reshape2)
library(ggplot2)
library(readxl)
library(lubridate)
library(data.table)
library(Rmisc)
library(tidyfast)
library(Hmisc)

################# SET WORKING DIRECTORY
wd <- "/Users/vkumar/Box Sync/LAB/PSY todo/2023-01-15 JCMS Sens Analysis/Data/"
setwd(wd)

################# READ INPUT FILE 
filename <- "/Users/vkumar/Box Sync/LAB/PSY todo/2023-01-15 JCMS Sens Analysis/Data/2023-01-20 TOM_PSYQuery2.txt"

d1 <- fread(filename, sep = '\t')

################# CREATE A FOLDER FOR OUTPUT FILES 
#  Create an empty folder in directory that is one level higher than the input file

dir.create(file.path("../PSY_Processed"), showWarnings = FALSE)

```

# Filtering data to find the right animals.
Filtering done by 
1. Dose 
2. Tester
3. Strain

```{r}
################# FILTER DATA AND PROCESS DATA
################# CHANGE THESE AS NEEDED 

################# FILTER BY DOSE ###################################
Dose <- 10
d2 <- d1[DrugDose == Dose]
#################################################################### 


################# FILTER BY TESTER #################################
d2 <- d2[ExptDoneBy == 'Marina']

# this can be a list -- example
# d2 <- d1[ExptDoneBy %in% .('Sandeep','Marina')]
#################################################################### 

################# FILTER BY STRAIN #################################
d2 <- d2[strainName %in% .('C57BL/6NJ','C57BL/6J')]
#################################################################### 

```

## Use this code to determine which animals have not been tested 13 times, or have been tested more than 13 times. 

Outputs - 
1. "Filtered_PSY_Data_d2file.tsv" - d2 file. this is filtered file from above. Its been filtered for dose, tester, strain. It has NOT been filtered for number of tests or for number of tests. 

*This part is important.* There are cases of mice with duplicated data.
Use the following two table that are produced to clean data.

2. "StrainFrequencyTable.csv" - this will have a frequency table for each
strain

3. "Number_tests.csv" - this table has one row for each mouse the number of
times its been tested.

**IMPORTANT TO LOOK AT THESE TABLES AND DETERMINE IF ALL DATA IS ACCURATE** 


```{r, warning=FALSE}

################# SORT BY MouseID AND TestDate
d2 <- d2[order(MouseID, TestDate)]

################# ADD TEST NUMBER AS A NEW COLUMN
d2 <- d2[ , TestNumber := order(TestDate), by = MouseID]

################# CREATE SOME OUTPUTS FOR SANITY CHECKS
# DATA WILL BE IN THE PSY_Processed DIRECTORY
setwd(wd)
setwd("../PSY_Processed")

################# WRITE d2
fwrite(d2, file = "Filtered_PSY_Data_d2file.tsv", sep = "\t")

################# WRITE FREQUENCY OF TESTS 
# This can be used to find animals that have not not been tested 13 times for instance.

# Frewuency table that counts will list MouseID and frequency

freqTable <- d2[, .N ,by = .(MouseID, strainName)] 
freqTable <- freqTable[order(strainName, N)] #orders the table above

# condenced table by strain, also renames N to numberTests

StrainFreqTable <- freqTable[, .N, by = .(strainName, numberTests = N)]  
fwrite(StrainFreqTable, "StrainFrequencyTable.csv")

# create and write a table by MouseID to determine which animals are missing data 

fwrite(d2[, .N ,by = .(MouseID, strainName)], 
       file = "Number_tests.csv", 
       sep = ",")


```
## Filter data to find the set of animals to analyze AND Sanity Check Calculations and Plots

This part will only operate on mice that have been tested a certain
number of times.

A completed Sens series should have 13 total tests.

need to remove all empty rows still

FileOutputs (usually csv files)

"d3" is a filter of "d2" that has 13 tests. 

1. "AnimalsWith13Tests.csv" is an output of d3

2. "LongFormatOf_d3.csv" - is the long format of "d3"

3. "ByAnimalCompiled_PSYdata*.pdf" - Minute by minute data for each animal. All test dates are plotted on one page. I print the comment for each group here. 

4. "ByAnimalCompiled_PSYdata_B_*.pdf" - Minute by minute data for each animal and each date. One trace per page. So a large file. I also print the comments for each animal here. 

```{r}
################# PLOT ALL TRACES FOR EACH ANIMAL

################# Only work with animals that have been tested a certain number of times

numberTestsNeeded <- 13 # required number of tests

MiceToKeep <- d2[, .N, by = MouseID][, MouseID[N >= numberTestsNeeded]]

# filter the mice with at 13 tests
d3 <- d2[MouseID %in% MiceToKeep]


# rearrange columns 
### I am rearranging the columns. move all minute by minute data to the end
d3 <- d3 %>% relocate(num_range("",0:120), .after = TestNumber)
d3 <- d3 %>% relocate(num_range("-",60:1), .after = TestNumber)


# write a table of animals with 13 tests
setwd(wd)
setwd("../PSY_Processed")
fwrite(d3, "AnimalsWith13Tests.csv", sep = ",")

# make a long table for plotting
file4 <- melt(d3, id = 1:69)
fwrite(file4, "LongFormatOf_d3.csv", sep = ",")

# Making the date correct, if the date is not formatted correctly, adjust this I changed "%m/%d/%Y"
file4$TestDate <- mdy_hms(file4$TestDate)  

# as.integer causes some weird oscillation in data. as.numeric is the way to go.  
file4$variable <- as.numeric(as.character(file4$variable))

##########################################################
### Plot By AnimalID - one plot for each animal
##########################################################

toPlotAnimalNames <- levels(as.factor(file4$AnimalName))

outFileNamePDF <- paste("../PSY_Processed/ByAnimalCompiled_PSYdata",format(Sys.time(),"%Y-%m-%d(%H_%M%p)"),".pdf", sep = "") ## .. means one directory level up

pdf(outFileNamePDF, 8,4)

for (toPlot in toPlotAnimalNames) {
# extract the animal to plot
  toPlot1 <- file4[file4$AnimalName == toPlot,]

# plot the animal
  p1 <- ggplot(data = toPlot1, aes(x = variable, y = value, group = TestDate, color = as.factor(TestDate))) +
    geom_line() + 
    annotate("text", x = 0, y  = 90, label= toPlot1$GroupComments, color = "red") +
    coord_cartesian(ylim = c(1,100)) +
    geom_vline(xintercept = 0, color = 'blue', linewidth = 0.2) +
    ggtitle(toPlot) 

  print(p1)
}

dev.off()

##########################################################
### Plot By AnimalID - one plot for each test
##########################################################

outFileNamePDF <- paste("../PSY_Processed/ByAnimalCompiled_PSYdata_B_",format(Sys.time(),"%Y-%m-%d(%H_%M%p)"),".pdf", sep = "") ## .. means one directory level up

toPlotAnimalNames <- levels(as.factor(file4$AnimalName))

pdf(outFileNamePDF, 4,2)

for (toPlot in toPlotAnimalNames) {
# extract the animal to plot
  toPlot1 <- file4[file4$AnimalName == toPlot,]

  testDates <- levels(as.factor(toPlot1$TestDate))

for (toPlot2 in testDates) {
    toPlot3 <- toPlot1[as.factor(toPlot1$TestDate) == toPlot2,]

# plot the animal
  p1 <- ggplot(data = toPlot3, aes (x = variable, y = value)) +
    geom_line() + annotate("text", x = 0, y  = 90, label= toPlot3$IndividualComments, color = "red") +
    coord_cartesian(ylim = c(1,100)) +
    geom_vline(xintercept = 0, color = 'blue', linewidth = .2) +
    ggtitle(paste(toPlot, toPlot2))

  print(p1)
  }
  }

dev.off()

```

## Calculate minute by minute average of activity for each Strain

This section will produce the following summary files. They can be used in Excel for further analysis. 

CSV files

1.    "Min_SummaryPlots*.csv" - this has strain, test number, minute by minute avg 

2.    SEM_Strain_TestNo*.csv - this has strain, test number, minute by minute SEM data

Plots

"Min_SummaryPlots\*.pdf" - Has 2 plots.

1. minute by minute for each day of sens for each strain 

2. minute by minute for each animal faceted by strain.

```{r}
setwd(wd)

##################### EXPORT SUMMARY TABLES

# these are column names, from minutes -60 to 120 will be used in calculatons. 
# This can be changed as needed for other columns. I 'm using a trick to get a sequence of numbers.
# We could easily use c("column1", "Column 10")

cols <- as.character(seq(from = -60, to = 120, by = 1))

# MEAN
d5 <- d3[,lapply(.SD, mean, na.rm=T), by = .(strainName, TestNumber),
.SDcols = cols] []

# SEM
d6 <- d3[,lapply(.SD, function(x) sd(x, na.rm=T)/sqrt(length(na.omit(x)))), by = .(strainName, TestNumber),
.SDcols = cols] []

fwrite(d5,paste("../PSY_Processed/Mean_Strain_TestNo_",Dose, format(Sys.time(),"mg_%Y-%m-%d(%H_%M%p)"),".csv", sep = ""))
fwrite(d6,paste("../PSY_Processed/SEM_Strain_TestNo_",Dose, format(Sys.time(),"mg_%Y-%m-%d(%H_%M%p)"),".csv", sep = ""))

 
##################### PLOTS minute by minute by strain
p2 <- ggplot(data = file4, mapping = aes(x = variable, y = value, color = as.factor(strainName), fill = as.factor(strainName))) + 
  theme_bw() +
  stat_summary(fun.data = mean_se, geom = "ribbon", alpha = 0.3) +
  stat_summary(fun = mean, geom="line", color = "grey50", linewidth = 0.2) +
  facet_grid(TestNumber~.) +
  xlim(-60, 60) +
  geom_vline(xintercept = 0, color = 'red', linewidth = .2) + 
  ggtitle(paste("Sens 10mg/kg - Per strain")) +
  theme(legend.position = "top") + theme(legend.title=element_blank()) + # removes legend title
  labs(x = "Time (Min)", y = "Distance (cm)") +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Test Number (1-4 Saline | 5-10 Cocaine | 11 Saline | 12, 13 Cocaine)", breaks = NULL, labels = NULL)) 

# PLOT minute by minute by each animals
p3 <- ggplot(data = file4, mapping = aes(x = variable, y = value, color = as.factor(strainName), fill = as.factor(strainName), group=AnimalName)) + 
  theme_bw() +
  geom_line(linewidth = 0.2) +
  facet_grid(TestNumber~strainName) +
  xlim(-60, 60) +
  geom_vline(xintercept = 0, color = 'red', linewidth = 0.2) +
  ggtitle(paste("Sens 10mg/kg - Per Animal")) +
  theme(legend.position = "top") + theme(legend.title=element_blank()) + # removes legend title
  labs(x = "Time (Min)", y = "Distance (cm)") +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Test Number (1-4 Saline | 5-10 Cocaine | 11 Saline | 12, 13 Cocaine)", breaks = NULL, labels = NULL)) 
# using secondary axis option from ggplot2 "https://stackoverflow.com/questions/11353287/how-do-you-add-a-general-label-to-facets-in-ggplot2"


outFileNamePDF <- paste("../PSY_Processed/Min_SummaryPlots",format(Sys.time(),"%Y-%m-%d(%H_%M%p)"),".pdf", sep = "") 
## .. means one directory level up
pdf(outFileNamePDF, 4, 10)

print(p2)
print(p3)

dev.off()

```

## Plot the Sens Data.

I'm recycling code from previous Sens Analysis. Using File d3 from above
chunk Plots will appear in "Sensitization Analysis\*.pdf" file

```{r}
setwd(wd)
######################################################################
# START PDF STATEMENT
# 
#######################################################################
outFileNamePDF <- paste("../PSY_Processed/Sensitization Analysis-",format(Sys.time(),"%Y-%m-%d(%H_%M%p)"),".pdf", sep = "") 
## .. means one directory level up
pdf(outFileNamePDF, 10,8)

######################################################################
# sum5minavg 
# 
#######################################################################

# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum5minavg, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum5minavg, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum5minavg, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = sum5minavg, colour = as.factor(AnimalName)), data=d3, shape = 21,size = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = sum5minavg, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = sum5minavg, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = sum5minavg, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)

######################################################################
# sum60min 
# 
#######################################################################

# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum60min, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum60min, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = sum60min, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom")+ theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = sum60min, colour = as.factor(AnimalName)), data=d3, shape = 21,linewidth = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = sum60min, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = sum60min, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = sum60min, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)


######################################################################
# NetResponse 
# 
#######################################################################

# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = NetResponse, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = NetResponse, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = NetResponse, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = NetResponse, colour = as.factor(AnimalName)), data=d3, shape = 21,size = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = NetResponse, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = NetResponse, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = NetResponse, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)

######################################################################
# AvgResp30min 
# 
#######################################################################

# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = as.factor(AnimalName)), data=d3, shape = 21,size = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = AvgResp30min, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = AvgResp30min, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = AvgResp30min, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)

######################################################################
# BaselineAvg 
# 
#######################################################################

# PLOT OF MEAN AND SEM WITH EACH TEST POINT
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = strainName,group = strainName), data=d3) +
  geom_point(shape = 21,size = 2) + 
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT OF MEAN AND SEM FOR EACH STRAIN, NO INDIVIDUAL POINTS
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se")
print(sp1)

# PLOT FACET BY STRAIN, POINT FOR EACH TEST
sp1 <- ggplot(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = strainName,group = strainName), data=d3) +
  theme_bw() + theme(legend.position = "top") +
  geom_point(shape = 21,size = 2) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(linewidth = 1.0, stat = "summary") +
  stat_summary(fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) 
print(sp1)


# PLOT FACET BY STRAIN, EACH ANIMAL'S ACTIVITY IS PLOTTED FOR EACH DAY
sp1 <- ggplot() +
  theme_bw() + theme(legend.position = "bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5)) +
  geom_point(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = as.factor(AnimalName)), data=d3, shape = 21,size = 2) +
  geom_line(aes(x = as.factor(TestNumber),y = BaselineAvg, group = strainName), data=d3, linewidth = 1.0, stat = "summary") +
  stat_summary(aes(x = as.factor(TestNumber),y = BaselineAvg, group = strainName), data=d3, fun.data = "mean_se") +
  facet_grid(facets = . ~ strainName) +
  geom_line(aes(x = as.factor(TestNumber),y = BaselineAvg, colour = as.factor(AnimalName), group = as.factor(AnimalName)), data = d3)

print(sp1)

######################################################################
# END PDF STATEMENT
# 
dev.off()
#######################################################################

```

data.table notes <https://rdatatable.gitlab.io/data.table/index.html>

good tutorial
<https://raw.githack.com/uo-ec510-2020-spring/lectures/master/05-datatable/05-datatable.html#1>
